Software variability in fMRI analysis:\\ comparing between-tool and numerical errors

### 2021-12-03 paper/paper.tex
Updates in bibliography, minor edits in methods

\author{Ali Salari$^1$, Other Authors$^1$, Tristan Glatard$^1$ \\
$^1$ Department of Computer-Science and Software Engineering, Concordia University, Montreal, Canada}

@@

To validate our instrumentations, we first verified that non-instrumented
executions of the same pipeline on the same dataset led to identical
results \tristan{did you fix random seeds and if so in which pipelines?}. We also listed the pipeline
library dependencies using the \texttt{ldd} Linux utility and verified that
(1) the tested pipelines were dynamically linked to the GNU libmath library, and 
(2) there was no alternative implementation of elementary mathematical functions in the pipeline dependencies.
Finally, we verified that the use of Fuzzy libmath affected computational results.

@@

We measured within-tool (WT) variability by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error, the minimal distance between two consecutive floating-point numbers. The resulting samples are equally plausible estimates of
the true numerical result. Moreover, to evaluate numerical perturbations at different magnitudes, 
we repeated the FSL analyses for virtual
precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
and double-precision values. We also repeated the FSL analyses for virtual 
precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having 
set the virtual precision to $t=24$~bits for single-precision values.

We evaluated WT and BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps, by computing the
standard deviation of t-statistic maps across pipelines (BT) or Fuzzy Libmath samples (WT). For BT, we computed the standard deviation 
across a given pair of tools.
For WT, we computed the standard deviation across the 3 
Fuzzy Libmath samples of a given tool. Moreover, we computed WT variability for a pair of tools (A, B) as follows:
\begin{equation}
  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
  \label{eq:wt-pair}
\end{equation}
where $\sigma_{WT(.)}$ is the within-tool variability for a given tool.

>>>>

\author{Ali Salari$^1$, Yohan Chatelain$^1$, [other co-authors], Gregory Kiar$^2$, Tristan Glatard$^1$ \\
$^1$ Department of Computer-Science and Software Engineering, Concordia University, Montreal, Canada\\
$^2$Center for the Developing Brain, Child Mind Institute, New York, NY, USA}

@@

In~\cite{salari2021accurate}, Fuzzy Libmath was shown to accurately
simulate the effect of Linux operating system updates in structural
pre-processing pipelines. To validate our pipeline instrumentations for the present study, we first verified that non-instrumented
executions of the same pipeline on the same dataset led to identical
results \tristan{did you fix random seeds and if so in which pipelines?}. We also listed the pipeline
library dependencies using the \texttt{ldd} Linux utility and verified that
(1) the tested pipelines were dynamically linked to the GNU libmath library, and 
(2) there was no alternative implementation of elementary mathematical functions in the pipeline dependencies.
Finally, we verified that the use of Fuzzy libmath affected computational results.

@@

We measured numerical (within-tool -- WT) variability by running the same analyses three
times using Fuzzy Libmath with a virtual precision of $t=53$~bits for
double-precision values and $t=24$~bits for single-precision values. These
values were chosen such that the numerical perturbation simulates machine
error. The resulting samples are equally plausible estimates of
the true numerical result at the precision used by the pipelines. Moreover, to evaluate numerical perturbations at different magnitudes, 
we repeated the FSL analyses for virtual
precisions ranging from $t=1$~bit to $t=24$~bits for single-precision
and double-precision values. We also repeated the FSL analyses for virtual 
precisions ranging from $t=24$~bits to $t=53$~bits for double-precision values, having 
set the virtual precision to $t=24$~bits for single-precision values.

We evaluated WT and BT variability for thresholded as well as unthresholded
group-level and subject-level t-statistics maps, by computing the
standard deviation of t-statistic maps across pipelines (BT) or Fuzzy Libmath samples (WT). For BT, we computed the standard deviation 
across a given pair of tools.
For WT, we computed the standard deviation across the three
Fuzzy Libmath samples of a given tool. Moreover, we computed WT variability for a pair of tools (A, B) as follows:
\begin{equation}
  \sigma_{WT(A,B)}^2 = \sigma_{WT(A)}^2 + \sigma_{WT(B)}^2,
  \label{eq:wt-pair}
\end{equation}
where $\sigma_{WT(.)}$ is the numerical variability for a given tool
and $\sigma_{WT(., .)}$ is the numerical variability for a pair of tools.
